Evaluation

### 2018-01-22 evaluation.Rmd
move everything to eval page and fine tune

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(dplyr)
library(portalr)
source("forecast_tools.R")
```

@@

```{r, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=9, fig.height=15}

@@

species_of_interest = c('BA','DM','DO','PP','OT','NA')
species_names = portalr::FullPath('PortalData/Rodents/Portal_rodent_species.csv', '~') %>%
  read.csv(stringsAsFactors=FALSE) %>%
  select(species = speciescode, full_species_name = scientificname)

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(dplyr)
library(portalr)
source("tools/forecast_tools.R")

species_of_interest = c('BA','DM','DO','PP','OT','NA','total')
species_names = portalr::FullPath('PortalData/Rodents/Portal_rodent_species.csv', '~') %>%
  read.csv(stringsAsFactors=FALSE) %>%
  select(species = speciescode, full_species_name = scientificname)

## How have the models recently?

This graph shows the forecast error of some recent forecasts for the most recent sampling dates. This is shown for all model types in the columns and several common rodent species (and total rodents) in the rows. These forecasts are for the entire site. 
```{r recent_eval, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=15, fig.height=15}
library(ggrepel)
# The length of time to look backward
prior_new_moons = 12

new_moon_file = portalr::FullPath('PortalData/Rodents/moon_dates.csv', '~')
new_moons = read.csv(new_moon_file)
most_recent_newmoon = max(new_moons$newmoonnumber)
earliest_newmoon_to_graph = most_recent_newmoon - prior_new_moons

observation_data = species_abundance %>%
  bind_rows(total_abundance) %>% 
  filter(newmoonnumber %in% earliest_newmoon_to_graph:most_recent_newmoon)

#Get the all the forecasts made during observation period
forecast_data = compile_forecasts(use_hindcasts = FALSE) %>%
  filter(newmoonnumber %in% earliest_newmoon_to_graph:most_recent_newmoon)

# Round so that large decimal place differences do not cause
# similar forecasts to be seen as distinct. 
forecast_data$estimate = round(forecast_data$estimate, 3)
forecast_data$LowerPI = round(forecast_data$LowerPI, 3)
forecast_data$UpperPI = round(forecast_data$UpperPI, 3)

# Keep 1 forecast per unique initial_newmoon, when there are multiple forecasts
# using the same initial_newmoon, use the first one.
forecast_dates_to_keep = forecast_data %>%
  select(date, initial_newmoon) %>%
  distinct() %>%
  group_by(initial_newmoon) %>%
  filter(date == min(date)) %>%
  ungroup() %>%
  mutate(keep='yes')

forecast_data = forecast_data %>%
  left_join(forecast_dates_to_keep, by=c('date','initial_newmoon')) %>%
  filter(keep=='yes') %>%
  select(-keep)
##################################

forecast_errors = forecast_data %>%
  left_join(observation_data, by=c('species','level','currency','newmoonnumber')) %>% 
  mutate(rmse = sqrt((estimate - actual)^2))

# Sometimes there are model runs on the same day and with all the same info,
# this gets rid of those
forecast_errors = forecast_errors %>%
  distinct()

# Drop any entries that don't have an observation
# (ie. a rainy sample period)
forecast_errors = forecast_errors %>%
  filter(!is.na(actual))

# Filter to common species and apply the full name
forecast_errors = forecast_errors %>%
  filter(species %in% species_of_interest) %>%
  left_join(species_names, by='species') %>%
  select(-species) %>%
  rename(species=full_species_name)

# Add a newline to the longer species names
forecast_errors$species = with(forecast_errors, ifelse(nchar(species)>15, stringr::str_replace(species,' ','\n'), species))

###################################
forecast_date_labels = forecast_errors %>%
  group_by(date, model, species, censusdate) %>%
  summarize(y_position = mean(rmse)) %>%
  ungroup() %>%
  distinct() %>%
  filter(!is.na(censusdate)) %>%
  group_by(date, model, species) %>%
  filter(as.Date(censusdate) == min(as.Date(censusdate))) %>%
  ungroup() %>%
  distinct() %>%
  mutate(label_text = paste0('Issued: ',date))

####################################
# Setup a nice palette for when prior issue dates become greater than 9
getPalette = colorRampPalette(RColorBrewer::brewer.pal(9, "Set1"))
#Randomize pallette so nearby regions are not similar colors
set.seed(2)
large_color_palette = sample(getPalette(40))
#####################################
ggplot(forecast_errors, aes(x=censusdate, y=rmse, group=as.character(date), color=as.character(date))) +
  geom_point(size=4)+
  geom_line(size=1.5) +
  # label individual lines if there gets to be a lot of them. remove the legend if this is turned on.
  #geom_label_repel(data=forecast_date_labels, aes(label=label_text, y=y_position),
  #                 fill='grey90',size=4,nudge_y = 40, force=5,
  #                 segment.alpha = 0.3, segment.size = 0.8) +
  scale_color_brewer(palette='Dark2') +
  #scale_color_manual(values=large_color_palette) +
  facet_grid(species~model) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=0.9, debug = FALSE),
        axis.text = element_text(size=12),
        axis.title = element_text(size=20),
        strip.text = element_text(size=12),
        legend.text = element_text(size=15),
        legend.title = element_text(size=18)) +
  labs(x='Sample Date',y='RMSE', color='Forecast Issue\nDate')

@@

```{r hindcast_eval, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=9, fig.height=15}

@@
